Assertion Contracts

### 2024-06-27 papers/assertion-contract/assertion_contract_app/paper.tex
Some formulas are added

\begin{itemize}
\item There are two ways to reward validators to encourage their participation: to discover a counterexample, and to provide computational proof that they have executed the program to search for a counterexample, even if none is found.
\item To prioritize the discovery of counterexamples, the reward for finding a counterexample is set significantly higher than that for providing a computational proof.
\item The computation of the negation formula is integrated into the computational proof to ensure that the validators cannot produce the proof without executing a negation check.
\item Because the rewards may be limited in quantity, validators are motivated to swiftly verify the parameters to claim the reward before others do so.
\end{itemize}
\subsubsection{Computation Proof}
The next step is to devise a computation proof to incentivize computational effort by utilizing the proof of work concept. We use a target hash $t$ that sets the required difficulty level to earn the reward. The objective of the validator is to generate a computation that produces a hash smaller than or equal to the target hash $t$.

The key problem is how to reproduce the computation proof off-chain to on-chain. To this end, a seed $s$ is introduced in each verification process. Let $n$ denote the size of domain $A$. Each number $i$ in the range [1, $n$] corresponds to an element in $A$. The seed $s$ then reduces modulo $n$ to obtain $i$, which is  associated with an element $a$ in $A$.

The sequence of operations can be described by the following steps:
\begin{enumerate}
\item Generate a random seed $s$: 
\begin{align}
s \in \mathbb{N} 
\end{align}
where $s$ is a random natural number.
\item Calculate $n$ as the size of the domain $A$: 
\begin{align}
n = |A|
\end{align}
where $\mid A \mid$ denotes the number of elements in the domain $A$.
\item Determine $i \in [1, n]$ using the modulo operation:
\begin{align}
i = (s \mod n) + 1
\end{align}
\item Find the corresponding element $a$ in $A$:
\begin{align}
a=A[i]
\end{align}
where $A[i]$ denotes the $i$-th element in $A$.
\item Calculate the hash \( t_s \) corresponding to the seed \( s \):
\begin{align}
t_s = \text{hash}(s, \text{eval}(\neg P_{a}))
\end{align}
where \(\text{hash}\) is the hash function and \(\text{eval}\) is the function that returns the significant components of the evaluation of the negation \(\neg P_{a}\).

For example, it returns \(a[i]\) and \(a[i + 1]\) in the evaluation of the negation predicate \(a[i] < a[i+1]\) in the sorted array example.
\end{enumerate}

@@

%\section{Domain Specification Language for Assertion}
%\section{Converter - Generating an Assertion Contracts} 
\section{Prototype Implementations}
\note{TODO}
\subsection{Sorted array}
%There are two possible data structures for storing such records: map and array. For example, if we assume that a caller can submit only one parameter at a time, we can use a map that associates the caller's address with the parameter's hash and timestamp.
%\begin{lstlisting}[numbers=none]
%   struct Parameter {
%        bytes32 p_hash;
%        uint256 timestamp;
%    }
% mapping (address => Parameter) public parameters;
%\end{lstlisting}

>>>>

\begin{itemize}
\item There are two ways to reward validators to encourage their participation: to discover a counterexample, and to provide computational proof that they have executed the program to search for a counterexample, even if none is found.
\item To prioritize the discovery of counterexamples, the reward for finding a counterexample is set significantly higher than that for providing a computational proof.
\item The computation of the negation formula is integrated into the computational proof to ensure that the validators cannot produce the proof without executing a negation check.
\item Because the rewards may be limited in quantity, validators are motivated to swiftly verify the parameters to claim the reward before others do so.
\end{itemize}
\subsubsection{Computation Proof}
The next step is to devise a computation proof to incentivize computational effort by utilizing the proof-of-work concept. We introduce a target hash $t$ that sets the required difficulty level to earn the reward. The objective of the validator is to generate a computation that produces a hash smaller than or equal to the target hash $t$.

A key challenge is reproducing the computation proof off-chain to on-chain. To address this, we introduce a seed $s$ in each verification process. Let $n$ denote the size of domain $A$. Each number $i$ in the range $[1, n]$ corresponds to an element in $A$. The seed $s$ is reduced modulo $n$ to obtain $i$, which is associated with an element $a$ in $A$.
\paragraph{Estimate the number of seeds:} We first estimate the number of seeds needed to find a counterexample with a given expected probability.

Let's assume the unsatisfied element is only at the first element $a_0$ in $A$, the size of the domain is $n$, and the number of seeds is $m$. Each seed independently has a probability of $1/n$ to detect the problem and a probability of $\frac{n-1}{n}$ not to detect the problem. Hence, if we assume that each seed associated with an element $a$ in $A$ is chosen independently from a uniform distribution, the probability that no validator checks the element $a_0$ converges to $0$ as the number of seeds approaches infinity:

\[
  \lim_{m \to \infty} \frac{(n-1)^m}{n^m} = \lim_{m \to \infty} \left( \frac{n-1}{n} \right)^m = 0
\]

Depending on the expected probability to detect the problem, the number of seeds to generate can be determined. Assume that $p$ is the probability of finding at least one counterexample. To calculate the number of seeds $m$ needed to achieve a given expected probability $p$ of finding a counterexample, we can use the complementary probability. We have:
\begin{itemize}
\item The probability of a single seed failing to find a counterexample is $\frac{n-1}{n}$.
\item The probability of all $m$ seeds failing to find a counterexample is $\left(\frac{n-1}{n}\right)^m$.
\item Therefore, the probability of at least one of the $m$ seeds finding a counterexample is:
\[
p = 1 - \left(\frac{n-1}{n}\right)^m
\]
\end{itemize}

Rearranging to solve for \( m \):

\[
\left(\frac{n-1}{n}\right)^m = 1 - p
\]

Taking the natural logarithm on both sides:

\[
\ln\left(\left(\frac{n-1}{n}\right)^m\right) = \ln(1 - p)
\]

\[
m \ln\left(\frac{n-1}{n}\right) = \ln(1 - p)
\]

Solving for \( m \):

\[
m = \frac{\ln(1 - p)}{\ln\left(\frac{n-1}{n}\right)}
\]

%For example, given \( n = 10000 \) and \( p = 0.8 \), the number of seeds \( m \) needed to achieve the desired probability is calculated as follows:
%\[
%m = \frac{\ln(1 - 0.8)}{\ln\left(1 - \frac{1}{10000}\right)}
%\]

%\[
%m = \frac{\ln(0.2)}{\ln\left(1 - 0.0001\right)}
%\]

%\[
%m = \frac{-1.6094}{\ln(0.9999)}
%\]

%\[
%m = \frac{-1.6094}{-0.0001}
%\]

%\[
%m \approx 16094
%\]

%Therefore, approximately \( 16094 \) seeds are needed to have an \( 80\% \) chance of finding a counterexample.

\paragraph{Set up the taget hash:}
The above seed number $m$ can then be used to set up the target hash $t$. The target hash $t$ can be calculated based on a desired difficulty level. In proof-of-work systems, difficulty is typically adjusted to ensure that the average efforts to find a valid hash. The target hash $t$ is often calculated by taking the maximum possible hash value and dividing it by the difficulty level:
\[
t = \frac{H_{\text{max}}}{D}
\]

Where:
\begin{itemize}
  \item \( H_{\text{max}} \) is the maximum possible hash value.
  \item \( D \) is the difficulty level, which can be adjusted based on the expected number of trials.
\end{itemize}

Given the expected number of seeds to generate, 
the difficulty level $D$ can be adjusted accordingly.  bTypically, $D$ is set to ensure that a valid hash is found approximately once every $m$ trials.

If we want the probality of finding a valid hash to be $\frac{1}{m}$, then:
\[
D=m
\]

For a 256-bit hash, $H_{\text{max}}$ is $2^{256} - 1 $.  using the maximum hash value and the difficulty level, the target hash $t$ is:

\[
\frac{2^{256} - 1}{m}
\]

\paragraph{Find a seed:}
The sequence of operations to find a seed can be described by the following steps:
\begin{enumerate}
\item Generate a random seed $s$: 
\begin{align}
s \in \mathbb{N} 
\end{align}
where $s$ is a random natural number.
\item Calculate $n$ as the size of the domain $A$: 
\begin{align}
n = |A|
\end{align}
where $\mid A \mid$ denotes the number of elements in the domain $A$.
\item Determine $i \in [1, n]$ using the modulo operation:
\begin{align}
i = (s \mod n) + 1
\end{align}
\item Find the corresponding element $a$ in $A$:
\begin{align}
a=A[i]
\end{align}
where $A[i]$ denotes the $i$-th element in $A$.
\item Perform the negation check $\neg P_{a}$ to detect a counterexample if any.

\item Calculate the hash \( t_s \) corresponding to the seed \( s \):
\begin{align}
t_s = \text{hash}(s, \text{eval}(\neg P_{a}))
\end{align}
where \(\text{hash}\) is the hash function and \(\text{eval}\) is the function that returns the significant components of the evaluation of the negation \(\neg P_{a}\).

For example, it returns \(a[i]\) and \(a[i + 1]\) in the evaluation of the negation predicate \(a[i] < a[i+1]\) in the sorted array example.
\item Compare \( t_s \) to the target hash to determin whether it is a reward computation.
\end{enumerate}

@@

\section{Domain Specification Language for Assertion}
\note{To discuss: if it make sense to design}
\section{Converter - Generating an Assertion Contracts} 
\note{To discuss: if it make sense to implement}
\section{Prototype Implementations}
\note{TODO}
\subsection{Sorted array}
%There are two possible data structures for storing such records: map and array. For example, if we assume that a caller can submit only one parameter at a time, we can use a map that associates the caller's address with the parameter's hash and timestamp.
%\begin{lstlisting}[numbers=none]
%   struct Parameter {
%        bytes32 p_hash;
%        uint256 timestamp;
%    }
% mapping (address => Parameter) public parameters;
%\end{lstlisting}
